---
title: "NHANES data exploration"
author: "Swen Kuh"
date: "2022-08-18"
output: html_document
---
We run into singularity issues for models lme1-lme4 (after removing pregnancy indicator (too many missing obs.), secondhand smoke exposure, potassium intake level, smoking tobacco/cigarette and sodium level), and the warning stops for model lme5. However, when using the urine subsample, we run into the same issue again. There is no warning issued after we remove the education variable. But should we? I didn't want to remove sodium level at first (I'm interested in the effects) but in the context of MRP, we are more concerned of predicting than inference of the effects? 

The data cleaning code is shown at the bottom of the document. 

## Running the models - lme4
```{r, warning=F, echo=F, error=F, message=F}
library(brms)
library(lme4)
library(tidyverse)
load(here::here('nhanes/nhanes_res.RData'))
```

```{r}
# # full model
# nhmodel_lme1 = glmer(high_bp ~ (1|age) + (1|ethnicity) + gender + (1|educ) +
#                        (1|marital_status) + phys_act + overweight + (1|dep_severity) +  
#                        (1|BMI_range) + (1|alc_exc) + (1|diabetes) + trb_sleep + 
#                        (1|pov_level) + smk_tobcig + smk2_exp +
#                        (1|sodium_lvl) + (1|potassium_lvl), data = nhdata, family="binomial")
# 
# # remove smk2_exp 
# nhmodel_lme2 = glmer(high_bp ~ (1|age) + (1|ethnicity) + gender + (1|educ) +
#                         (1|marital_status) + phys_act + overweight + (1|dep_severity) +  
#                         (1|BMI_range) + (1|alc_exc) + (1|diabetes) + trb_sleep + 
#                         (1|pov_level) + smk_tobcig  +
#                         (1|sodium_lvl) + (1|potassium_lvl), data = nhdata, family="binomial")
# 
# # remove potassium level
# nhmodel_lme3 = glmer(high_bp ~ (1|age) + (1|ethnicity) + gender + (1|educ) +
#                        (1|marital_status) + phys_act + overweight + (1|dep_severity) +  
#                        (1|BMI_range) + (1|alc_exc) + (1|diabetes) + trb_sleep + 
#                        (1|pov_level) + smk_tobcig  + (1|sodium_lvl) , data = nhdata, family="binomial")
# 
# # remove smk_tobcig
# nhmodel_lme4 = glmer(high_bp ~ (1|age) + (1|ethnicity) + gender + (1|educ) +
#                        (1|marital_status) + phys_act + overweight + (1|dep_severity) +  
#                        (1|BMI_range) + (1|alc_exc) + (1|diabetes) + trb_sleep + 
#                        (1|pov_level) + (1|sodium_lvl) , data = nhdata, family="binomial")

# remove sodium level - no more singularity error -  6786 observations (after na.omit) 
nhmodel_lme5 = glmer(high_bp ~ (1|age) + (1|ethnicity) + gender + (1|educ) +
                       (1|marital_status) + phys_act + overweight + (1|dep_severity) +  
                       (1|BMI_range) + (1|alc_exc) + (1|diabetes) + trb_sleep + 
                       (1|pov_level) , data = nhdata, family="binomial")

summary(nhmodel_lme5)
```



```{r}
# subsample - 2224 observations (after na.omit)
nhmodel_lme5_sub = glmer(high_bp ~ (1|age) + (1|ethnicity) + gender +
                        (1|marital_status) + phys_act + overweight + (1|dep_severity) +  
                       (1|BMI_range) + (1|alc_exc) + (1|diabetes) + trb_sleep + 
                       (1|pov_level) , data = nh_urn_sub, family="binomial")
summary(nhmodel_lme5_sub)

# MSE
( pred.lme = predict(nhmodel_lme5, type="response") %>% 
  as.data.frame(.) %>% 
  mutate(pred.bin = ifelse(. >= 0.5, 1, 0), 
         truth = as.numeric(nhsub$high_bp), 
         bias = pred.bin - truth) %>% 
  summarise(mse = mean(bias^2)) )
```

## Running the models - brms
```{r}
## brms ####
# nhmodel = brm(high_bp ~ (1|age) + (1|ethnicity) + gender + (1|educ) +
#                 (1|marital_status) + phys_act + overweight + (1|dep_severity) +  
#                 (1|BMI_range) + (1|alc_exc) + (1|diabetes) + trb_sleep + 
#                 (1|pov_level), 
#               data = nhdata, 
#               seed = 2345,
#               chain = 2,
#               backend = "cmdstanr",
#               family = bernoulli(link = "logit"), 
#               control = list(adapt_delta = 0.9)) 

# MSE 
( pred.brms = predict(nhmodel)[,1] %>% 
    as.data.frame(.) %>% 
    mutate(pred.bin = ifelse(. >= 0.5, 1, 0), 
           truth = as.numeric(nhsub$high_bp), 
           bias = pred.bin - truth) %>% 
    summarise(mse = mean(bias^2)) )

# nhmodel_urn = brm(high_bp ~ (1|age) + (1|ethnicity) + gender + (1|educ) + 
#                     (1|marital_status) + phys_act + overweight + (1|dep_severity) +  
#                     (1|BMI_range) + (1|alc_exc) + (1|diabetes) + (1|trb_sleep) + 
#                     (1|pov_level) , 
#                   data = dat2, 
#                   seed = 2345,
#                   backend = "cmdstanr",
#                   family = bernoulli(link = "logit"), 
#                   control = list(adapt_delta = 0.9)) 

# MSE
( pred.brms.urn = predict(nhmodel_urn)[,1] %>% 
    as.data.frame(.) %>% 
    mutate(pred.bin = ifelse(. >= 0.5, 1, 0), 
           truth = as.numeric(nh_urn_sub$high_bp), 
           bias = pred.bin - truth) %>% 
    summarise(mse = mean(bias^2)) )
```






## Data cleaning
```{r setup, include=FALSE}
library(here)
library(foreign)
library(tidyverse)
library(questionr) # recode.na()
library(GGally) # ggpairs()
library(forcats) #fct_relevel
# https://www.aihw.gov.au/reports/risk-factors/risk-factors-to-health/contents/high-blood-pressure
# https://www.mayoclinic.org/diseases-conditions/high-blood-pressure/symptoms-causes/syc-20373410
# family history, stress level, chronic condition, diabetes, sleep apnea

```


```{r nhanes, echo=F, warning=F, message=F, eval=F}
## 2017-2020 
# demographic - age, race, sex/gender, education, married status ####
demo1720 = read.xport(here('data/NHANES 2017-20/P_DEMO.XPT'))  %>% # 15560 obs total
  filter(RIDAGEYR >= 18) %>% # 10195 for 16 y/o and above, 9693 for 18 y/o 
  mutate(age = cut(RIDAGEYR, breaks = c(18, 29, 39, 49, 59, 69, 80), right = F, labels = F),
         age = recode(as.factor(age), '1' = '18-29', '2' = '30-39', '3' = '40-49', '4' = '50-59', '5' = '60-69', '6' = '70-80'),
         ethnicity = recode(as.factor(RIDRETH3), '1' = 'MexAme', '2' =  'OtherHisp', '3' = 'NonHispWhite', '4' = 'NonHispBlack', '6' = 'NonHispAsian', '7' = 'Other'),    #'1-Mexican American; 2-Other Hispanic; 3-Non-Hispanic White; 4-Non-Hispanic Black; 6-Non-Hispanic Asian; 7-Other Race including Multi-Racial'
         gender = recode(as.factor(RIAGENDR), '1' = 'male', '2' = 'female'),  # '1-male; 2-female
         DMDEDUC2 = recode(as.factor(DMDEDUC2), '7' = '9'), #1-Less than 9th grade; 2-9-11th grade (Includes 12th grade with no diploma);3-High school graduate/GED or equivalent;4-Some college or AA degree;5-College graduate or above; 7-Refused; 9-Don't Know,
         educ = recode(DMDEDUC2, '1' = 'less9th', '2' = '9-11thgrade', '3' = 'HSgrad', '4' = 'SomecollegeAA', '5' = 'Collegegrad', '9' = 'refused/dont know'),
         marital_status = recode(as.factor(DMDMARTZ), '77' = 'refused/dont know'),
         marital_status = recode(marital_status, '99' = 'refused/dont know', '1' = 'married/living tgt', '2' = 'widowed/separated', '3' = 'never married')) %>% 
  select(SEQN, age, ethnicity, gender, educ, marital_status)

# sum(demo1720$RIDAGEYR >= 18 & demo1720$RIDAGEYR < 25)
# sum(demo1720$RIDAGEYR >= 65 & demo1720$RIDAGEYR < 75)
# sum(demo1720$RIDAGEYR >= 75 & demo1720$RIDAGEYR < 80)
```

```{r, echo=F, warning=F, message=F, eval=F}
# outcome - high blood pressure 
bp = read.xport(here('data/NHANES 2017-20/P_BPQ.XPT')) %>% 
  mutate(high_bp = recode(as.factor(BPQ020), '9' = '0'), # recode '9 - don't know' as 'no'
         high_bp = recode(high_bp, '2' = '0'),
         high_bp = fct_relevel(high_bp, c('0','1'))) %>% 
  select(SEQN, high_bp) 

# covariates - physical activity 
paq = read.xport(here('data/NHANES 2017-20/P_PAQ.XPT')) %>%  # 18 y/0 and above
  mutate(phys_act = recode(as.factor(PAQ650), '2' = '0'), # '1 - yes to doing vigorous intensity for 10 mins'; '2 - no'
         phys_act = fct_relevel(phys_act, c('0','1'))) %>% 
  select(SEQN, phys_act) 

# covariates - overweight
mcq = read.xport(here('data/NHANES 2017-20/P_MCQ.XPT')) %>% 
  mutate(overweight = recode(as.factor(MCQ080), "9" = "2"),  # recode '9 - don't know' as no
         overweight = recode(overweight, '2' = '0'),
         overweight = fct_relevel(overweight, c('0','1'))) %>%
  select(SEQN, overweight)

# covariates - sodium (day1), potassium ####
# https://www.sciencedirect.com/science/article/pii/S000282239900423X?casa_token=oW171FSUo-YAAAAA:MMGugjEsy6PAlexkLyy2s8Z-rImUytgQA8rxNaTVEG80LaSY8gru1cH7R8_9n3vAwZQTcJm8Zfk
dr1 = read.xport(here('data/NHANES 2017-20/P_DR1TOT.XPT')) %>% 
  mutate(sodium_intake1 = replace_na(DR1TSODI, 0),
         potassium_intake1 = replace_na(DR1TPOTA,0)) %>% 
    select(SEQN, sodium_intake1, potassium_intake1) 

# covariates - sodium (day2), potassium ####
# https://academic.oup.com/advances/article/4/3/368S/4591617
dr2 = read.xport(here('data/NHANES 2017-20/P_DR2TOT.XPT')) %>% 
  mutate(sodium_intake2 = replace_na(DR2TSODI,0),
         potassium_intake2 = replace_na(DR2TPOTA,0)) %>%
  select(SEQN, sodium_intake2, potassium_intake2) 


# covariates - alcohol use ####
# https://www.sciencedirect.com/science/article/pii/S0168827816301817?casa_token=qFMfm2F1dfgAAAAA:u3isyHCDhaDfRXGGSwgqEgit3IiKRhQIX5rwtO234bVAv8tSJPbr3LaPBX3eE2zvNfTc9iq6RB8
# https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0218463
 # 1 = alcohol consumption between 1–14 (men) or 1–7 (women) standard drinks per week; 2 = alcohol consumption exceeding 14 drinks (men) or 7 drinks (women) per week. 
alc = read.xport(here('data/NHANES 2017-20/P_ALQ.XPT')) %>%  # 8965 obs - aged 12 and over sampled but data restricted to 18 and over
  left_join(., demo1720, by="SEQN") %>% 
  mutate(alcohol_use_day =  case_when(ALQ111 == '2' | ALQ121 == '0' ~ 0, # never drank alcohol and never in the last 12 months
                                      ALQ130 == "777" | ALQ130 == "999" ~ 999 , # combine refused and dont know
                                      ALQ130 == 1 | ALQ130 == 2 | ALQ130 == 3 | ALQ130 == 4 | ALQ130 == 5 
                                      | ALQ130 == 6 | ALQ130 == 7 | ALQ130 == 8 | ALQ130 == 9 | ALQ130 == 10
                                      | ALQ130 == 11 | ALQ130 == 12 | ALQ130 == 13 | ALQ130 == 14 | ALQ130 == 15 ~ ALQ130),
         alcohol_use_day = as.numeric(alcohol_use_day),
         alcohol_use_week = alcohol_use_day * 7,
         alc_exc = case_when(alcohol_use_week == 0 ~ 0,
                             gender == 'male' & alcohol_use_week >= 1 & alcohol_use_week <=14 ~ 1, 
                             gender == 'female' & alcohol_use_week >= 1 & alcohol_use_week <= 7 ~ 1, 
                             gender == 'male' & alcohol_use_week >14 ~ 2, 
                             gender == 'female' & alcohol_use_week >7 ~ 2), 
         alc_exc = recode(as.factor(alc_exc), '0' = 'no', '1' = 'moderate', '2' = 'excessive')) %>% # avg alcoholic drinks/day - past 12 mos
  select(SEQN, alc_exc) 

# covariates - bmi ####
bmi = read.xport(here('data/NHANES 2017-20/P_BMX.XPT')) %>% 
  rename(BMI_raw = BMXBMI) %>% 
  drop_na(BMI_raw) %>% 
  mutate(BMI = cut(BMI_raw, breaks = c(11.8, 18.5, 25, 30, 35, 40, 92.6), right = F, labels=F),
         BMI = as.factor(BMI), 
         BMI_range = recode(BMI, '1' = "underweight", '2' = 'healthy', '3' = 'overweight', '4' = 'obeseI', '5' = 'obeseII', '6' = 'obeseIII')) %>% 
  select(SEQN, BMI_range)
# str(bmi$BMI_range)


# covariates - depression ####
dep = read.xport(here('data/NHANES 2017-20/P_DPQ.XPT')) %>%  # same as alcohol use, aged 12 and over sampled but data restricted to 18 and over 
  mutate(DPQ010 = recode.na(as.factor(DPQ010), c('9', '7')),
         DPQ020 = recode.na(as.factor(DPQ020),  c('9', '7')),
         DPQ030 = recode.na(as.factor(DPQ030),  c('9', '7')),
         DPQ040 = recode.na(as.factor(DPQ040),  c('9', '7')),
         DPQ050 = recode.na(as.factor(DPQ050),  c('9', '7')),
         DPQ060 = recode.na(as.factor(DPQ060),  c('9', '7')),
         DPQ070 = recode.na(as.factor(DPQ070),  c('9', '7')),
         DPQ080 = recode.na(as.factor(DPQ080),  c('9', '7')),
         DPQ090 = recode.na(as.factor(DPQ090),  c('9', '7'))) %>% 
    mutate_if(is.factor, as.numeric) %>% 
         mutate(dep_score = DPQ010 + DPQ020 + DPQ030 + DPQ040 + DPQ050 +
                     DPQ060 + DPQ070 + DPQ080 + DPQ090,
                dep_level = cut(dep_score, breaks = c(1, 9, 14, 19, 36), right = T, label=F),
                dep_severity = recode(dep_level, '1' = "mild", '2' = "moderate", '3' = "moderately severe", '4' = "severe"),
                dep_severity = as.factor(dep_severity)) %>% 
  select(SEQN, dep_severity)
# summary(dep)
# sum(na.rm(dep$dep_score) >= 20 & na.rm(dep$dep_score) <= 27)
  
# covariates - diabetes ####
diab = read.xport(here('data/NHANES 2017-20/P_DIQ.XPT')) %>% 
  mutate(diabetes = case_when(DIQ010 == '9' | DIQ010 == '2'~ 0,
                              DIQ010 == '1' ~ 1,
                              DIQ010 == '3' ~ 2),
         diabetes = recode(as.factor(diabetes), '0' = 'no', '1' = 'yes', '2' = 'borderline')) %>%  # ever told have diabetes - 9 - don't know - code as no; 3 is borderline
  select(SEQN, diabetes)

# covariates - pregnancy ####
preg = read.xport(here('data/NHANES 2017-20/P_RHQ.XPT')) %>% # 5314 - female only - can impute 0 for men? 
  mutate(preg_fem = case_when(RHD143 == "2" ~ 0,
                              RHD143 == "1" ~ 1,
                              RHD143 == "9" ~ 999),
         preg_fem = as.factor(preg_fem)) %>%  # are you pregnant now - 9 - don't know - code as no
 select(SEQN, preg_fem)

# covariates - trouble sleeping ####
sleep = read.xport(here('data/NHANES 2017-20/P_SLQ.XPT')) %>% # Ever told doctor had trouble sleeping?
  mutate(trb_sleep = case_when(SLQ050 == "7" | SLQ050 == "9" | SLQ050 == "2" ~ 0,
                               SLQ050 == "1" ~ 1),
         trb_sleep = as.factor(trb_sleep)) %>%  # 7 - refused to answer; 9 - don't know
    select(SEQN, trb_sleep)

# covariates - tobacco ####
tob = read.xport(here('data/NHANES 2017-20/P_SMQRTU.XPT')) %>% # Smoked any tobacco product last 5 days? 
  mutate(smk_tob_last5 = case_when(SMDANY == '2' ~ 0, 
                                   SMDANY == '1' ~ 1),
         smk_tob_last5 = as.factor(smk_tob_last5)) %>% 
    select(SEQN, smk_tob_last5) 

# covariates - cigarette ####
cig = read.xport(here('data/NHANES 2017-20/P_SMQ.XPT')) %>% # Do you now smoke cigarettes or smoke at least 100 cigarettes in life?
  mutate(cig_now = case_when(SMQ040 == '1' | SMQ040 == '2' ~ 1,  # recode 'some days' as 'occasional' smoker = 1
                             SMQ040 == '3' | SMD030 == '0' | SMQ020 == '2' | SMQ621 == '1' ~ 0, 
                             SMQ020 == '7' | SMQ020 == '9' ~ 999),
         cig_now = as.factor(cig_now)) %>%  # recode 'not at all - 3' as '0' so '1' is smoke and '0' is doesn't smoke
  select(SEQN, cig_now)

# covariates - 2nd hand smoke indoors ####
smk2a = read.xport(here('data/NHANES 2017-20/P_SMQSHS.XPT')) %>% # Last 7-d someone smoked indoors 
  select(SEQN, SMQ858, SMQ862, SMQ868, SMQ872, SMQ876, SMQ880, SMQ940) %>%  # at work, restaurant, bar, car, other indoor areas, e-cig
  mutate(smk2_ind = case_when(SMQ858 == 1 | SMQ862 == 1 | SMQ868 == 1 | SMQ872 == 1 | 
                                SMQ876 == 1 | SMQ880 == 1 |SMQ940 == 1 ~ 1, 
                              SMQ858 == 2 | SMQ862 == 2 | SMQ868 == 2 | SMQ872 == 2 | 
                                SMQ876 == 2 | SMQ880 == 2 |SMQ940 == 2 ~ 0,
                              TRUE ~ NA_real_ ),
         smk2_ind = as.factor(smk2_ind)) %>% 
   select(SEQN, smk2_ind)  
# summary(smk2a)

# covariates - 2nd hand smoke at home ####
smk2b = read.xport(here('data/NHANES 2017-20/P_SMQFAM.XPT')) %>% # of people who live here smoke tobacco?
  mutate(smk2_home = case_when(SMD460 == '0' | SMD460 == '777' | SMD460 == '999' ~ '0', # 777 - refused, 999 - don't know
                               SMD460 == '1' | SMD460 == '2' ~ '1'),
         smk2_home = as.factor(smk2_home)) %>% # recode 2+ smokers to be same group as 1+
  select(SEQN, smk2_home)  

# covariates - poverty level ####
# https://ajph.aphapublications.org/doi/pdf/10.2105/AJPH.91.5.781
pov = read.xport(here('data/NHANES 2017-20/P_INQ.XPT')) %>%
  mutate(pov_level = as.factor(INDFMMPC),
         pov_level = recode(pov_level, '1' = 'low income', 
                            '2' = 'middle income', 
                            '3' = 'high income', 
                            '7' = 'refused',
                            '9' = 'dont know')) %>% 
  select(SEQN, pov_level)

# nhdata ####
fulldat = left_join(demo1720, bp, by='SEQN') %>% 
  left_join(., paq, by='SEQN') %>% 
  left_join(., mcq, by='SEQN') %>% 
  left_join(., dep, by='SEQN') %>% 
  left_join(., dr1, by='SEQN') %>% 
  left_join(., dr2, by='SEQN') %>% 
  left_join(., bmi, by='SEQN') %>% 
  left_join(., alc, by='SEQN') %>% 
  left_join(., diab, by='SEQN') %>% 
  left_join(., preg, by='SEQN') %>% 
  left_join(., sleep, by='SEQN') %>% 
  left_join(., tob, by='SEQN') %>% 
  left_join(., cig, by='SEQN') %>% 
  left_join(., smk2a, by='SEQN') %>% 
  left_join(., smk2b, by='SEQN') %>% 
  left_join(., pov, by='SEQN') # demo1720 - 6960; dat1 - 3489; dat2 - 2317

nhdata = fulldat %>% 
  mutate(smk2_exp = ifelse(smk2_ind == 1 | smk2_home == 1, 1, 0), # exposure to smoke at work or at home
         smk2_exp = as.factor(smk2_exp), 
         smk_tobcig = ifelse(smk_tob_last5 == 1 | cig_now == 1, 1, 0),  # smokes tobacco or cigarette 
         smk_tobcig = as.factor(smk_tobcig),  
         preg_yes = case_when(preg_fem == "999" ~ 999,
                              gender == 'male' | preg_fem == '0' ~ 0,
                              preg_fem == "1" ~ 1), # if male then no for preg
         preg_yes = recode(as.factor(preg_yes), '999' = 'dont know'), 
         sodium_avg = ifelse(sodium_intake1 == 0, sodium_intake2, ifelse(sodium_intake2 == 0, sodium_intake1, ((sodium_intake1 + sodium_intake2)/2) )), 
         potassium_avg = ifelse(potassium_intake1 == 0, potassium_intake2, ifelse(potassium_intake2 == 0, potassium_intake1, ((potassium_intake1 + potassium_intake2)/2))),
         sodium_lvl = case_when(is.na(sodium_avg) ~ NA_real_,
                                sodium_avg <= 1150 ~ 1,
                                sodium_avg > 1150 & sodium_avg <= 2300 ~ 2, 
                                sodium_avg > 2300 & sodium_avg <= 3450 ~ 3,
                                sodium_avg > 3450 ~ 4),
         sodium_lvl = recode(as.factor(sodium_lvl), '1' = 'low', '2' = 'intermediate', '3' = 'high', '4' = 'very high'),
         potassium_lvl = case_when(is.na(potassium_avg) ~ NA_real_,
                                   potassium_avg <= 2400 ~ 1, 
                                   potassium_avg > 2400 & potassium_avg <= 3600 ~ 2,
                                   potassium_avg > 3600 ~ 3),
         potassium_lvl = recode(as.factor(potassium_lvl), '1' = 'low', '2' = 'adequate','3' = 'high')) %>% 
  select(-c(smk2_ind, smk2_home, smk_tob_last5, cig_now, preg_fem, sodium_intake1, sodium_intake2, potassium_intake1, potassium_intake2))

summary(nhdata)
dim(nhdata)
```


```{r, eval=F}
# subsample
# insulin
ins = read.xport(here('data/NHANES 2017-20/P_INS.XPT'))

# urine sample
urn = read.xport(here('data/NHANES 2017-20/P_UM.XPT'))

nh_ins = merge(ins, nhdata, by='SEQN') # 4438 obs
nh_urn = merge(urn, nhdata, by='SEQN') # 2988 obs


```

## Exploratory plots

```{r, echo=F, warning=F, error=F, message=F}
# summary(glm(high_bp~ age+gender+ethnicity+educ+diabetes+trb_sleep+smk2_exp + smk_tobcig + alc_exc, data=nhdata, family='binomial'))

ggpairs(nhdata[,c(7,2:6,8:9)])
ggpairs(nhdata[,c(7,10:19)])
plot(BMI_range~high_bp, main="BMI~high_BP", data = nhdata)
```